untitled

### 2016-04-23 model.tex
refining

\begin{align}
  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0,\\
  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0. 
\end{align}  
These criteria are applicable to any adiabatic fluid when $s$
is the usual definition of entropy. For the dusty problem, we may take 
$s=\ln{(1-\epsilon)}+\mathrm{const.}$ from the definition of $s$ and  
the equation of state (Eq. \ref{eos}) with uniform sound-speed.

@@

The stability conditions are thus
\begin{align}
  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\epsilon &> 0,  \\
  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\epsilon}{\p
    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \epsilon}{\p r} \right) & > 0. 
\end{align}
In accretion disks, $\rhog$, $\rhod$ and $\epsilon$ are expected to
decrease outwards and away from the midplane, 
\begin{align*}  
  \frac{\p\rhog}{\p |z|} < 0, \quad \frac{\p \rhog}{\p r} < 0,   
\end{align*}
and similarly for $\rhod$ and $\epsilon$. Thus 
$\nabla\rhog\cdot\nabla\epsilon>0$, so the first criterion is
typically satisifed.

To evaulate the second criterion, we can consider $z>0$ without loss
of generality. Then, inserting the vertical shear profile, 
Eq. \ref{vshear}, we find the second criterion becomes

\begin{align} 
  \kappa^2\left|\frac{\p\epsilon}{\p z}\right| + c_s^2\left|\frac{\p
    \epsilon}{\p r}\frac{\p\epsilon}{\p z}\frac{\p\ln{\rho}}{\p
    r}\right| - c_s^2 \left(\frac{\p\epsilon}{\p
    r}\right)^2\left|\frac{\p\ln{\rho}}{\p z}\right| > 0.  \label{criterion2}
\end{align}

This equation shows that the vertical gradient in dust-fraction (and
thus the dust-to-gas ratio $\rhod/\rhog$) has a stabilizing effect for
axisymmetric perturbations. Instead, the only way to destabilize the
system is a sufficiently large radial gradient in the dust-fraction
and/or vertical (total) density gradient (the last term).

Note that Eq. \ref{criterion2} is only applicable for the assume disk
structure described above: decreasing dust and gas densities, as well
as the dust-fraction, away from the star and away from the
midplane. For less typical disk structures, one must evalulate above the
critera seperately.

>>>>

\begin{align}
  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0,\\
  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, 
\end{align}  
where $\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the
square of the epicylic frequency. We assume $\kappa^2>0$. 
These criteria are applicable to any adiabatic fluid when $s$
is the usual definition of entropy. For the dusty problem, we may take 
$s=\ln{(1-\epsilon)}+\mathrm{const.}$ from the definition of $s$ and  
the equation of state (Eq. \ref{eos}) with uniform sound-speed.

@@

The stability conditions are thus
\begin{align}\label{dusty_solberg}
  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\epsilon &> 0,  \\
  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\epsilon}{\p
    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \epsilon}{\p r} \right) & > 0. 
\end{align}

In accretion disks, $\rhog$, $\rhod$ and $\epsilon$ are expected to
decrease away from the mid-plane
\begin{align*}  
  \frac{\p\rhog}{\p |z|} < 0,
%, \quad \frac{\p \rhog}{\p r} < 0,   
\end{align*}
and similarly for $\rhod$, $\epsilon$, and the dust-to-gas ratio $\tepsilon$. 
We may thus evaluate the stability criteria at $z>0$ without loss of
generality. We insert the vertical shear profile,
Eq. \ref{vshear}, into Eq. \ref{dusty_solberg} and write the result 
in terms of $\tepsilon$, to find

\begin{align} 
  &1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1+
    \tepsilon\right)^2}\frac{\Omega^2}{\kappa^2}\left(\frac{\p\ln{\rhog}}{\p\ln{r}}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}} + r^2\frac{\p\ln{\rhog}}{\p z}
  \frac{\p\ln{\tepsilon}}{\p z} \right)>0,   \\
&-\frac{\p\ln{\tepsilon}}{\p z} \notag\\
  &- \frac{\tepsilon
  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}\frac{\Omega^2}{\kappa^2}
\frac{\p\ln{\tepsilon}}{\p\ln{r}}\left( \frac{\p\ln{\rhog}}{\p\ln{r}}\frac{\p\ln{\tepsilon}}{\p
    z} -\frac{\p\ln{\rhog}}{\p z}\frac{\p\ln{\tepsilon}}{\p\ln{
    r}}\right) > 0,
%  \kappa^2\left|\frac{\p\epsilon}{\p z}\right| + c_s^2\left|\frac{\p
%    \epsilon}{\p r}\frac{\p\epsilon}{\p z}\frac{\p\ln{\rho}}{\p
%    r}\right| - c_s^2 \left(\frac{\p\epsilon}{\p
%    r}\right)^2\left|\frac{\p\ln{\rho}}{\p z}\right| > 0.  \label{criterion2}
\end{align}
as the stability criteria, where 
\begin{align}
  h_\mathrm{g}\equiv c_s/\left(r\Omega\right)
\end{align}
is the characteristic aspect-ratio of the gas component.

%This equation shows that the vertical gradient in dust-fraction (and
%thus the dust-to-gas ratio $\rhod/\rhog$) has a stabilizing effect for
%axisymmetric perturbations. Instead, the only way to destabilize the
%system is a sufficiently large radial gradient in the dust-fraction
%and/or vertical (total) density gradient (the last term).

%Note that Eq. \ref{criterion2} is only applicable for the assume disk
%structure described above: decreasing dust and gas densities, as well
%as the dust-fraction, away from the star and away from the
%midplane. For less typical disk structures, one must evalulate above the
%critera seperately.

\begin{align} 
  &1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1+
    \tepsilon\right)^2}\frac{\p\ln{\rhog}}{\p\ln{r}}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}} +
  \tepsilon\frac{z^2}{\widetilde{H}^2}>0,   \\ 
&1 - \frac{\tepsilon
  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right) > 0,
\end{align}
